Place of death in NRS COVID-19 deaths

### 2021-05-14 covid_nrs_place_of_death_visualisations.Rmd
added a plot of deaths over time by place, "normalised" so that the historic mean is set to 0 and only deviations are plotted.
added a plot of causes of death from 2020 until the latest available data.

@@

## Table of annual deaths by cause

@@

>>>>

### Place of death, deviation from mean

```{r}
(figure_place_of_death_deviation_from_mean <-
  merged_deaths_overall %>%
  filter(sex == "all" & age == "all") %>%
  filter(!place_of_death %in% c("All","Other")) %>%
  mutate(place_of_death = factor(place_of_death, levels = c("Hospital", "Home & other non-institution", "Care home"))) %>%  # needs to be factor with same levels, otherwise it messes up level order when added as geom_text later!)
  pivot_longer(cols = c(deaths_all_causes, deaths_non_covid), 
               names_to = "deaths_type", values_to = "deaths") %>%
  mutate(
    deaths = deaths - deaths_mean,
    deaths_min = deaths_min - deaths_mean,
    deaths_max = deaths_max - deaths_mean
  ) %>%
  ggplot(aes(x = date_w_c)) +
  scale_x_date(date_labels = "%d-%m-%y", date_breaks = "8 weeks", limits = c(earliest_date, most_recent_date_available_2021)) +
  annotate(geom = "rect", xmin = compute_start_date_from_week_number(week_number = 1, year_number = 2021), xmax = compute_start_date_from_week_number(week_number = weeks_available_2021, year_number = 2021), ymin = -Inf, ymax = Inf, fill = "grey90", alpha = 0.2) +  # shade area - geom_rect doesn't work here
  annotate(geom = "text", x=ymd("2021-01-07"), y=-320, label = "2021", hjust = 0, colour = "grey40") +
  geom_ribbon(aes(ymin = deaths_min, ymax = deaths_max), fill = "#4477aa", alpha = 0.5) +
  geom_line(aes(y = deaths, colour = deaths_type)) +
  geom_line(aes(y = 0), linetype = "dashed", colour="grey20") +
  facet_grid(~place_of_death) +
  scale_colour_manual(values = c("red","blue"),  # custom legend for the two lines
                      labels = c("All deaths", 
                                 "Non Covid-19 deaths"),
                      guide = guide_legend(label.hjust = 0.5, override.aes = list(size = 5))
                      ) +
  theme(
    legend.position = "top",
    plot.caption = element_text(size = 10, colour = "gray60"),
    plot.margin = margin(5, 15, 5, 5),
    axis.text.x = element_text(angle = 60, colour = "gray40", vjust=1, hjust=1)
  ) +
  labs(x = "Date w/c",
       y = "Weekly Deaths: difference from 2015-2019 mean (N)",
       subtitle = paste0(c(date_range_text), collapse="\n"),
       caption = paste0(c(death_registrations_text,removed_other_category_text,source_nrs_text), collapse="\n"),
       colour = "Cause of death"
       ) +
  NULL
)

save_output_file(filename = paste0(dir_outputs,"/figure_place_of_death_deviation_from_mean"), extensions = c(".pdf",".svg",".png"), plot = figure_place_of_death_deviation_from_mean, width = 28, height = 14, dpi = 300, units = "cm")
```

@@

## Table of 2020 annual deaths by cause

@@

## Deaths by cause, from 2020 until current day

```{r}
(figure_deaths_by_cause <-
  merged_deaths_cause_of_death %>%
  filter(cause_of_death!="All" & !place_of_death %in% c("Other","All")) %>%
  mutate(place_of_death = factor(place_of_death, levels = c("Hospital", "Home & other non-institution", "Care home"))) %>%  # needs to be factor with same levels, otherwise it messes up level order when added as geom_text later!)
  mutate(
   cause_of_death = factor(cause_of_death %>% str_wrap(., width = 15), levels = order_of_causes_of_death %>% str_wrap(., width = 15)),  # needed to allow converting to a number for plotting the line segments; str_wrap breaks the text up into lines
   cause_of_death = fct_rev(cause_of_death)
 ) %>%
  group_by(place_of_death, cause_of_death) %>%
  summarise(
    deaths = sum(number_of_deaths),
    deaths_mean = sum(deaths_mean),
    .groups="drop"
  ) %>%
  ggplot() +
  geom_col(aes(x = cause_of_death, fill = cause_of_death, y = deaths), alpha = 0.8, colour = "grey10") +
  geom_segment(aes(x = as.numeric(cause_of_death)-0.45, xend = as.numeric(cause_of_death)+0.45, y = deaths_mean, yend = deaths_mean), size = 0.8, colour = "grey1") +
  coord_flip() +
  scale_fill_viridis_d() +
  facet_wrap(~place_of_death) +
  scale_y_continuous(labels = function(x) format(x, big.mark = ",")) +
  theme(
    legend.position = "none", 
    plot.caption = element_text(size = 10, colour = "gray60"),
    legend.title = element_blank(), 
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank()
    ) +
  labs(
    x = NULL,
    y = "Deaths",
    subtitle = paste0(c(paste0("Bars are deaths ",str_remove_all(date_range_text,"Data ")),"Line segments are mean annual deaths 2015-2019"),collapse="\n")
  ) +
  NULL
)

save_output_file(filename = paste0(dir_outputs,"/figure_deaths_by_cause"), extensions = c(".pdf",".svg",".png"), plot = figure_deaths_by_cause, width = 10, height = 3, dpi = 300, units = "in")
```
